---
title: "SVCFit Downstream Evaluation (A2)"
output: html_document
---

## Parameters — edit these before running

```{r parameters}
# Portable paths. Resolve relative to this script's
# own location (Phylogeny_benchmark/scripts/..). Override with NPJ_PKG_ROOT, or run
# via evaluate_downstream.R, which rewrites this chunk from command-line arguments.
.fa   <- grep("^--file=", commandArgs(trailingOnly = FALSE), value = TRUE)
.sd   <- if (length(.fa)) normalizePath(dirname(sub("^--file=", "", .fa[1]))) else getwd()
.PB   <- Sys.getenv("NPJ_PB_ROOT", unset = normalizePath(file.path(.sd, ".."), mustWork = FALSE))
if (!dir.exists(file.path(.PB, "bootstrap"))) .PB <- normalizePath("..", mustWork = FALSE)
work_dir   <- file.path(.PB, "bootstrap")
input_dir  <- work_dir
truth_dir  <- file.path(.PB, "input_data", "hack")
out_dir    <- file.path(.PB, "output", "boots0_4")

cov        <- 50L
n_boots    <- 5L            # boots 0..4: boot=0 uses bare exp*/ dirs, b1–b4 use b{n}/exp*/
boot       <- NA_integer_   # set to an integer to restrict to one bootstrap index
expected_cases <- 100L
expected_no_tree <- NA_integer_
expected_correct_topologies <- NA_integer_
purities <- c(20L, 40L, 60L, 80L)
experiments <- paste0("exp", 1:5)
pre_stage  <- "pre_BAT"
post_stage <- "on_BAT"

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(lme4)
```

## Helper functions

```{r helpers}
true_clone_fracs_s1 <- function(pur) {
  idiv <- function(a, b) as.integer(a) %/% as.integer(b)
  s1_t1_c2 <- idiv(pur * 5L, 6L); s1_t1_c3 <- pur - s1_t1_c2
  s1_t2_c2 <- idiv(pur * 1L, 6L); s1_t2_c3 <- pur - s1_t2_c2
  list(
    t1 = c(trunk = 1.0, sub1 = s1_t1_c2 / pur, sub2 = s1_t1_c3 / pur),
    t2 = c(trunk = 1.0, sub1 = s1_t2_c2 / pur, sub2 = s1_t2_c3 / pur)
  )
}

load_truth_beds <- function(truth_dir, overlap = FALSE) {
  prefix <- if (overlap) c("c11", "c22", "c33") else c("c1", "c2", "c3")

  c1 <- read_delim(file.path(truth_dir, paste0(prefix[1], ".bed")), delim = "\t",
                   col_names = c("chr", "start", "end", "type", "info", "flank"),
                   show_col_types = FALSE) %>%
    mutate(true_clone = "truncal")

  c2 <- read_delim(file.path(truth_dir, paste0(prefix[2], ".bed")), delim = "\t",
                   col_names = c("chr", "start", "end", "type", "info", "flank"),
                   show_col_types = FALSE)

  c3 <- read_delim(file.path(truth_dir, paste0(prefix[3], ".bed")), delim = "\t",
                   col_names = c("chr", "start", "end", "type", "info", "flank"),
                   show_col_types = FALSE)

  sub1_private <- anti_join(c2, c1, by = c("chr", "start", "end")) %>%
    mutate(true_clone = "sub1")
  sub2_private <- anti_join(c3, c1, by = c("chr", "start", "end")) %>%
    mutate(true_clone = "sub2")

  bind_rows(c1, sub1_private, sub2_private) %>%
    select(chr, start, end, type, true_clone) %>%
    distinct()
}

wilson_ci <- function(k, n, conf = 0.95) {
  z     <- qnorm((1 + conf) / 2)
  p     <- k / n
  denom <- 1 + z^2 / n
  ctr   <- (p + z^2 / (2 * n)) / denom
  marg  <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / denom
  c(lower = max(0, ctr - marg), upper = min(1, ctr + marg))
}

normal_ci <- function(x, conf = 0.95) {
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 2) return(c(lower = NA_real_, upper = NA_real_))
  t_crit <- qt((1 + conf) / 2, df = n - 1)
  se     <- sd(x) / sqrt(n)
  c(lower = mean(x) - t_crit * se, upper = mean(x) + t_crit * se)
}

parse_facet_bed <- function(facet_file) {
  if (!file.exists(facet_file)) return(NULL)
  df <- read_tsv(facet_file, show_col_types = FALSE) %>%
    mutate(chrom = paste0("chr", chrom)) %>%
    select(chrom, start, end, tcn = `tcn.em`, lcn = `lcn.em`)
  df
}

parse_vcf_ao <- function(vcf_file) {
  lines       <- readLines(vcf_file, warn = FALSE)
  header_line <- lines[startsWith(lines, "#CHROM")][1]
  data_lines  <- lines[!startsWith(lines, "#")]

  if (length(data_lines) == 0 || is.na(header_line))
    return(data.frame(sv_id = character(0), AO = integer(0),
                      stringsAsFactors = FALSE))

  col_names <- strsplit(header_line, "\t")[[1]]
  fmt_idx   <- which(col_names == "FORMAT")
  samp_idx  <- fmt_idx + 1L

  rows  <- strsplit(data_lines, "\t")
  sv_id <- vapply(rows, `[[`, character(1), 3L)
  fmt   <- strsplit(vapply(rows, `[[`, character(1), fmt_idx),  ":")
  samp  <- strsplit(vapply(rows, `[[`, character(1), samp_idx), ":")

  ao_val <- mapply(function(f, s) {
    idx <- match("AO", f)
    if (is.na(idx) || idx > length(s)) return(NA_integer_)
    suppressWarnings(as.integer(s[[idx]]))
  }, fmt, samp, SIMPLIFY = TRUE)

  data.frame(sv_id = sv_id, AO = ao_val, stringsAsFactors = FALSE)
}

validate_tree_centroids <- function(tree_edges, centroids, case_id,
                                    lineage_thresh = 0.2,
                                    sum_thresh = 0.2) {
  edges <- as.data.frame(tree_edges, stringsAsFactors = FALSE)
  if (!all(c("parent", "child") %in% names(edges)) || nrow(edges) == 0L)
    stop(case_id, ": tree has no parent/child edges")
  edges$parent <- as.character(edges$parent)
  edges$child  <- as.character(edges$child)
  if (anyNA(edges$parent) || anyNA(edges$child) || any(edges$parent == edges$child))
    stop(case_id, ": tree contains missing or self-referential edges")
  if (sum(edges$parent == "root") != 1L)
    stop(case_id, ": tree must contain exactly one edge from root")
  if (anyDuplicated(edges$child))
    stop(case_id, ": a tree node has more than one parent")
  if (anyDuplicated(centroids$cluster_num))
    stop(case_id, ": centroid table contains duplicate cluster_num values")

  cluster_labels <- as.character(centroids$cluster_num)
  node_to_index <- function(node) {
    node <- as.character(node)
    direct <- match(node, cluster_labels)
    prefixed <- match(paste0("cluster", node), cluster_labels)
    numeric <- suppressWarnings(as.integer(node))
    result <- ifelse(!is.na(direct), direct,
                     ifelse(!is.na(prefixed), prefixed, numeric))
    as.integer(result)
  }

  tree_nodes <- unique(c(edges$parent[edges$parent != "root"], edges$child))
  node_indices <- node_to_index(tree_nodes)
  if (anyNA(node_indices) || any(node_indices < 1L | node_indices > nrow(centroids)))
    stop(case_id, ": tree references a cluster absent from cluster_centroids.csv")
  if (!setequal(node_indices, seq_len(nrow(centroids))))
    stop(case_id, ": tree and centroid table describe different cluster sets")

  # Confirm the stored edges still satisfy the constraints used by build_tree().
  ccf <- as.matrix(centroids[, c("t1_ccf", "t2_ccf")])
  storage.mode(ccf) <- "double"
  if (any(!is.finite(ccf))) stop(case_id, ": centroid CCFs are not finite")
  scoring_ccf <- ccf
  scoring_ccf[which.max(rowSums(scoring_ccf)), ] <- 1

  for (i in seq_len(nrow(edges))) {
    child_idx <- node_to_index(edges$child[i])
    parent_ccf <- if (edges$parent[i] == "root") rep(1, ncol(ccf)) else {
      parent_idx <- node_to_index(edges$parent[i])
      ccf[parent_idx, ]
    }
    if (any(ccf[child_idx, ] - parent_ccf > lineage_thresh + 1e-12))
      stop(case_id, ": stored tree violates lineage precedence")
  }
  for (parent in unique(edges$parent)) {
    child_idx <- node_to_index(edges$child[edges$parent == parent])
    child_sum <- colSums(scoring_ccf[child_idx, , drop = FALSE])
    parent_ccf <- if (parent == "root") rep(1, ncol(ccf)) else scoring_ccf[node_to_index(parent), ]
    if (any(child_sum - parent_ccf > sum_thresh + 1e-12))
      stop(case_id, ": stored tree violates the sum condition")
  }

  # Every node must be reachable from root; the unique-parent check then also
  # excludes disconnected cycles.
  reached <- "root"
  repeat {
    next_nodes <- unique(edges$child[edges$parent %in% reached])
    updated <- unique(c(reached, next_nodes))
    if (length(updated) == length(reached)) break
    reached <- updated
  }
  if (!all(tree_nodes %in% reached)) stop(case_id, ": tree contains unreachable nodes")
  invisible(TRUE)
}

eval_classification <- function(sv_stage, truth_svs, trunk_cluster) {
  matched <- sv_stage %>%
    rowwise() %>%
    mutate(
      truth_idx = {
        dists <- abs(truth_svs$start - POS)
        idx   <- which(truth_svs$chr == CHROM & dists < 500)
        if (length(idx) > 0) idx[which.min(dists[idx])] else NA_integer_
      },
      true_clone = if (!is.na(truth_idx)) truth_svs$true_clone[truth_idx] else NA_character_,
      true_clonality = case_when(
        true_clone == "truncal"            ~ "truncal",
        true_clone %in% c("sub1", "sub2") ~ "subclonal",
        TRUE                               ~ NA_character_
      )
    ) %>%
    ungroup() %>%
    filter(!is.na(true_clonality))

  if (nrow(matched) == 0)
    return(list(bal_acc = NA_real_, f1 = NA_real_, n_matched = 0L,
                sensitivity = NA_real_, specificity = NA_real_))

  tp <- sum(matched$inferred_clonality == "truncal"   & matched$true_clonality == "truncal",   na.rm = TRUE)
  fp <- sum(matched$inferred_clonality == "truncal"   & matched$true_clonality == "subclonal", na.rm = TRUE)
  fn <- sum(matched$inferred_clonality == "subclonal" & matched$true_clonality == "truncal",   na.rm = TRUE)
  tn <- sum(matched$inferred_clonality == "subclonal" & matched$true_clonality == "subclonal", na.rm = TRUE)

  sensitivity <- if (tp + fn > 0) tp / (tp + fn) else NA_real_
  specificity <- if (tn + fp > 0) tn / (tn + fp) else NA_real_
  bal_acc     <- if (!is.na(sensitivity) && !is.na(specificity))
                   (sensitivity + specificity) / 2 else NA_real_
  precision   <- if (tp + fp > 0) tp / (tp + fp) else NA_real_
  f1          <- if (!is.na(precision) && !is.na(sensitivity) && (precision + sensitivity) > 0)
                   2 * precision * sensitivity / (precision + sensitivity) else NA_real_

  # Sensitivity (truncal recall) and specificity (subclonal recall) are the two components
  # plotted in Supplementary Fig. S8D, returned and summarised per purity.
  list(bal_acc = bal_acc, f1 = f1, n_matched = nrow(matched),
       sensitivity = sensitivity, specificity = specificity)
}
```

## Main evaluation loop

```{r main-loop}
exps     <- experiments
boots    <- if (!is.na(boot)) boot else seq(0L, n_boots - 1L)

results        <- list()
sv_detail_list <- list()

for (b in boots) {
  for (pur in purities) {

    true_ccfs <- true_clone_fracs_s1(pur)

    for (exp_name in exps) {

      longi_dir <- if (b == 0L) {
        file.path(work_dir, "S1", paste0("c", cov, "p", pur), exp_name)
      } else {
        file.path(work_dir, "S1", paste0("c", cov, "p", pur),
                  paste0("b", b), exp_name)
      }
      source_longi_dir <- if (b == 0L) {
        file.path(input_dir, "S1", paste0("c", cov, "p", pur), exp_name)
      } else {
        file.path(input_dir, "S1", paste0("c", cov, "p", pur),
                  paste0("b", b), exp_name)
      }
      case_id <- sprintf("boot=%d S1/p%d/%s", b, pur, exp_name)

      cluster_file   <- file.path(longi_dir, "clustering", "sv2cluster.csv")
      centroids_file <- file.path(longi_dir, "clustering", "cluster_centroids.csv")
      tree_file      <- file.path(longi_dir, "tree", "tree_result.rds")
      no_tree_file   <- file.path(longi_dir, "tree", "NO_TREE.txt")

      required_files <- c(cluster_file, centroids_file)
      missing_files <- required_files[!file.exists(required_files) |
                                      file.info(required_files)$size <= 0]
      if (length(missing_files))
        stop(case_id, ": missing clustering output(s): ", paste(missing_files, collapse = ", "))
      has_tree <- file.exists(tree_file) && file.info(tree_file)$size > 0
      has_no_tree <- file.exists(no_tree_file) && file.info(no_tree_file)$size > 0
      if (has_tree == has_no_tree)
        stop(case_id, ": expected exactly one of tree_result.rds or NO_TREE.txt")

      tree_outcome <- if (has_tree) "tree" else "no_tree"
      cat(sprintf("  Evaluating %s (%s)...\n", case_id, tree_outcome))

      truth_svs  <- load_truth_beds(truth_dir, overlap = (exp_name != "exp1"))
      sv2cluster <- read_csv(cluster_file,   show_col_types = FALSE)
      centroids  <- read_csv(centroids_file, show_col_types = FALSE) %>%
        filter(pair == 1) %>%
        select(cluster_num, t1_ccf = f_pre, t2_ccf = f_day85)
      if (has_tree) {
        tree_res   <- readRDS(tree_file)
        tree_edges <- tree_res[[1]]
        stored_mcf <- as.matrix(tree_res[[2]])
        current_mcf <- as.matrix(centroids[, c("t1_ccf", "t2_ccf")])
        storage.mode(stored_mcf) <- "double"
        storage.mode(current_mcf) <- "double"
        if (!identical(dim(stored_mcf), dim(current_mcf)) ||
            !isTRUE(all.equal(unname(stored_mcf), unname(current_mcf),
                              tolerance = 1e-12, check.attributes = FALSE)))
          stop(case_id, ": tree_result.rds was built from different centroids")
        validate_tree_centroids(tree_edges, centroids, case_id)
      } else {
        tree_edges <- data.frame(parent = character(), child = character())
      }

      trunk_idx     <- tree_edges$child[tree_edges$parent == "root"]
      trunk_cluster <- if (length(trunk_idx) == 1)
        paste0("cluster", trunk_idx) else NA_character_

      sv2cluster <- sv2cluster %>%
        mutate(inferred_clonality = ifelse(cluster_num == trunk_cluster,
                                           "truncal", "subclonal"))

      class_pre  <- eval_classification(
        sv2cluster %>% filter(stage == pre_stage),  truth_svs, trunk_cluster)
      class_post <- eval_classification(
        sv2cluster %>% filter(stage == post_stage), truth_svs, trunk_cluster)

      non_trunk <- centroids %>% filter(cluster_num != trunk_cluster)

      ccf_inf_t1 <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      ccf_inf_t2 <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      best_asgn  <- NULL

      if (!is.na(trunk_cluster) && nrow(non_trunk) == 2) {
        assignments <- list(
          c(non_trunk$cluster_num[1], non_trunk$cluster_num[2]),
          c(non_trunk$cluster_num[2], non_trunk$cluster_num[1])
        )
        best_err  <- Inf
        best_t1   <- best_t2 <- NULL
        best_asgn <- NULL

        for (asgn in assignments) {
          inf_t1 <- c(
            trunk = centroids$t1_ccf[centroids$cluster_num == trunk_cluster][1],
            sub1  = centroids$t1_ccf[centroids$cluster_num == asgn[1]][1],
            sub2  = centroids$t1_ccf[centroids$cluster_num == asgn[2]][1]
          )
          inf_t2 <- c(
            trunk = centroids$t2_ccf[centroids$cluster_num == trunk_cluster][1],
            sub1  = centroids$t2_ccf[centroids$cluster_num == asgn[1]][1],
            sub2  = centroids$t2_ccf[centroids$cluster_num == asgn[2]][1]
          )
          err <- sum(abs(inf_t1 - true_ccfs$t1), na.rm = TRUE) +
                 sum(abs(inf_t2 - true_ccfs$t2), na.rm = TRUE)
          if (err < best_err) { best_err <- err; best_t1 <- inf_t1; best_t2 <- inf_t2; best_asgn <- asgn }
        }

        if (!is.null(best_t1)) {
          ccf_inf_t1 <- best_t1
          ccf_inf_t2 <- best_t2
        }
      }

      ccf_mae_t1 <- abs(ccf_inf_t1 - true_ccfs$t1)
      ccf_mae_t2 <- abs(ccf_inf_t2 - true_ccfs$t2)
      ccf_err_t1 <- true_ccfs$t1 - ccf_inf_t1
      ccf_err_t2 <- true_ccfs$t2 - ccf_inf_t2

      eff_cov_t1     <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      eff_cov_t2     <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      bpc_t1         <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      bpc_t2         <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      seg_len_t1     <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      seg_len_t2     <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      ascn_t1        <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      ascn_t2        <- setNames(rep(NA_real_, 3), c("trunk", "sub1", "sub2"))
      sv_detail_rows <- list()

      if (!is.na(trunk_cluster) && !is.null(best_asgn)) {
        clone_clust <- c(trunk = trunk_cluster,
                         sub1  = best_asgn[1], sub2 = best_asgn[2])
        for (tp_n in 1:2) {
          vcf_file   <- file.path(source_longi_dir, sprintf("t%d", tp_n), "svtyp",
                                  sprintf("svt_%s_t%d.vcf", exp_name, tp_n))
          facet_file <- file.path(source_longi_dir, sprintf("t%d", tp_n), "facet",
                                  sprintf("%s_t%d.bed", exp_name, tp_n))
          facet_df   <- parse_facet_bed(facet_file)
          stage_lbl  <- if (tp_n == 1L) pre_stage else post_stage
          tp_label   <- if (tp_n == 1L) "pre" else "post"

          sv_tp <- sv2cluster %>%
            filter(stage == stage_lbl) %>%
            mutate(sv_id = ifelse(!is.na(shared), shared,
                                  sub("^.*_(Manta.+)$", "\\1", event_id)))

          if (file.exists(vcf_file)) {
            ao_df <- parse_vcf_ao(vcf_file)
            sv_tp <- sv_tp %>% left_join(ao_df, by = "sv_id")
          } else {
            sv_tp <- sv_tp %>% mutate(AO = NA_integer_)
          }

          for (cl in c("trunk", "sub1", "sub2")) {
            sv_cl    <- sv_tp %>% filter(cluster_num == clone_clust[[cl]])
            ao_vals  <- sv_cl$AO[!is.na(sv_cl$AO)]
            len_vals <- abs(sv_cl$sv_len[!is.na(sv_cl$sv_len)])

            nd_flags <- if (!is.null(facet_df) && nrow(sv_cl) > 0L) {
              vapply(seq_len(nrow(sv_cl)), function(i) {
                chr <- sv_cl$CHROM[i]
                pos <- suppressWarnings(as.integer(sv_cl$POS[i]))
                idx <- which(facet_df$chrom == chr &
                             facet_df$start  <= pos & facet_df$end >= pos)
                if (length(idx) == 0L) NA_real_
                else as.numeric(facet_df$tcn[idx[1L]] != 2L |
                                facet_df$lcn[idx[1L]] != 1L)
              }, numeric(1))
            } else rep(NA_real_, nrow(sv_cl))

            mean_ao  <- if (length(ao_vals)  > 0L) mean(ao_vals)  else NA_real_
            mean_len <- if (length(len_vals) > 0L) mean(len_vals) else NA_real_
            bpc_val  <- as.numeric(nrow(sv_cl))
            ascn_val <- mean(nd_flags, na.rm = TRUE)

            if (tp_n == 1L) {
              eff_cov_t1[[cl]] <- mean_ao;  bpc_t1[[cl]] <- bpc_val
              seg_len_t1[[cl]] <- mean_len; ascn_t1[[cl]] <- ascn_val
            } else {
              eff_cov_t2[[cl]] <- mean_ao;  bpc_t2[[cl]] <- bpc_val
              seg_len_t2[[cl]] <- mean_len; ascn_t2[[cl]] <- ascn_val
            }

            if (nrow(sv_cl) > 0L)
              sv_detail_rows <- c(sv_detail_rows, list(tibble(
                boot          = b,
                purity        = pur,
                experiment    = exp_name,
                clone         = cl,
                timepoint     = tp_label,
                sv_type       = sv_cl$classification,
                ao            = sv_cl$AO,
                sv_len        = abs(sv_cl$sv_len),
                is_nondiploid = nd_flags
              )))
          }
        }
      }

      correct_topology <- FALSE
      topo_score       <- 0

      if (!is.null(tree_edges) && nrow(tree_edges) > 0) {
        root_children <- tree_edges$child[tree_edges$parent == "root"]
        if (length(root_children) == 1) {
          topo_score  <- topo_score + 1/3
          trunk_node  <- root_children[1]
          trunk_ch    <- tree_edges$child[tree_edges$parent == trunk_node]
          if (length(trunk_ch) == 2) {
            topo_score <- topo_score + 1/3
            leaf1 <- length(tree_edges$child[tree_edges$parent == trunk_ch[1]]) == 0
            leaf2 <- length(tree_edges$child[tree_edges$parent == trunk_ch[2]]) == 0
            if (leaf1 && leaf2) {
              topo_score       <- topo_score + 1/3
              correct_topology <- TRUE
            }
          }
        }
      }

      results <- c(results, list(tibble(
        boot               = b,
        purity             = pur,
        experiment         = exp_name,
        tree_outcome       = tree_outcome,
        n_svs_pre          = class_pre$n_matched,
        n_svs_post         = class_post$n_matched,
        n_clusters         = length(unique(sv2cluster$cluster_num)),
        bal_acc_pre        = class_pre$bal_acc,
        f1_pre             = class_pre$f1,
        bal_acc_post       = class_post$bal_acc,
        f1_post            = class_post$f1,
        sens_pre           = class_pre$sensitivity,
        spec_pre           = class_pre$specificity,
        sens_post          = class_post$sensitivity,
        spec_post          = class_post$specificity,
        ccf_inf_trunk_pre  = ccf_inf_t1["trunk"],
        ccf_inf_sub1_pre   = ccf_inf_t1["sub1"],
        ccf_inf_sub2_pre   = ccf_inf_t1["sub2"],
        ccf_inf_trunk_post = ccf_inf_t2["trunk"],
        ccf_inf_sub1_post  = ccf_inf_t2["sub1"],
        ccf_inf_sub2_post  = ccf_inf_t2["sub2"],
        ccf_mae_trunk_pre  = ccf_mae_t1["trunk"],
        ccf_mae_sub1_pre   = ccf_mae_t1["sub1"],
        ccf_mae_sub2_pre   = ccf_mae_t1["sub2"],
        ccf_mae_trunk_post = ccf_mae_t2["trunk"],
        ccf_mae_sub1_post  = ccf_mae_t2["sub1"],
        ccf_mae_sub2_post  = ccf_mae_t2["sub2"],
        ccf_err_trunk_pre  = ccf_err_t1["trunk"],
        ccf_err_sub1_pre   = ccf_err_t1["sub1"],
        ccf_err_sub2_pre   = ccf_err_t1["sub2"],
        ccf_err_trunk_post = ccf_err_t2["trunk"],
        ccf_err_sub1_post  = ccf_err_t2["sub1"],
        ccf_err_sub2_post  = ccf_err_t2["sub2"],
        eff_cov_trunk_pre  = eff_cov_t1["trunk"],
        eff_cov_sub1_pre   = eff_cov_t1["sub1"],
        eff_cov_sub2_pre   = eff_cov_t1["sub2"],
        eff_cov_trunk_post = eff_cov_t2["trunk"],
        eff_cov_sub1_post  = eff_cov_t2["sub1"],
        eff_cov_sub2_post  = eff_cov_t2["sub2"],
        bpc_trunk_pre      = bpc_t1["trunk"],
        bpc_sub1_pre       = bpc_t1["sub1"],
        bpc_sub2_pre       = bpc_t1["sub2"],
        bpc_trunk_post     = bpc_t2["trunk"],
        bpc_sub1_post      = bpc_t2["sub1"],
        bpc_sub2_post      = bpc_t2["sub2"],
        seg_len_trunk_pre  = seg_len_t1["trunk"],
        seg_len_sub1_pre   = seg_len_t1["sub1"],
        seg_len_sub2_pre   = seg_len_t1["sub2"],
        seg_len_trunk_post = seg_len_t2["trunk"],
        seg_len_sub1_post  = seg_len_t2["sub1"],
        seg_len_sub2_post  = seg_len_t2["sub2"],
        ascn_trunk_pre     = ascn_t1["trunk"],
        ascn_sub1_pre      = ascn_t1["sub1"],
        ascn_sub2_pre      = ascn_t1["sub2"],
        ascn_trunk_post    = ascn_t2["trunk"],
        ascn_sub1_post     = ascn_t2["sub1"],
        ascn_sub2_post     = ascn_t2["sub2"],
        correct_topology   = correct_topology,
        topo_score         = topo_score
      )))
      sv_detail_list <- c(sv_detail_list, sv_detail_rows)
    }
  }
}
```

## Compile results

```{r compile}
if (length(results) == 0) stop("No completed runs found.")

all_results <- bind_rows(results)
if (nrow(all_results) != expected_cases)
  stop(sprintf("Expected %d scored cases but found %d", expected_cases, nrow(all_results)))
n_no_tree <- sum(all_results$tree_outcome == "no_tree")
if (!is.na(expected_no_tree) && n_no_tree != expected_no_tree)
  stop(sprintf("Expected %d NO_TREE outcomes but found %d", expected_no_tree, n_no_tree))
n_correct_topologies <- sum(all_results$correct_topology)
if (!is.na(expected_correct_topologies) &&
    n_correct_topologies != expected_correct_topologies)
  stop(sprintf("Expected %d correct topologies but found %d",
               expected_correct_topologies, n_correct_topologies))
cat(sprintf("\n=== COMPLETENESS: n_expected=%d n_scored=%d n_no_tree=%d ===\n",
            expected_cases, nrow(all_results), n_no_tree))
write_csv(all_results, file.path(out_dir, "evaluation_summary.csv"))
cat(sprintf("\n=== SUMMARY (%d runs evaluated) ===\n", nrow(all_results)))

sv_detail_df <- if (length(sv_detail_list) > 0) bind_rows(sv_detail_list) else NULL
if (!is.null(sv_detail_df))
  write_csv(sv_detail_df, file.path(out_dir, "sv_detail_data.csv"))
```

## Classification summary

```{r classification}
class_summary <- all_results %>%
  group_by(purity) %>%
  summarise(
    n                 = n(),
    n_scored          = n(),
    n_no_tree         = sum(tree_outcome == "no_tree"),
    mean_bal_acc_pre  = mean(bal_acc_pre,  na.rm = TRUE),
    ci_lo_pre         = normal_ci(bal_acc_pre)["lower"],
    ci_hi_pre         = normal_ci(bal_acc_pre)["upper"],
    mean_bal_acc_post = mean(bal_acc_post, na.rm = TRUE),
    ci_lo_post        = normal_ci(bal_acc_post)["lower"],
    ci_hi_post        = normal_ci(bal_acc_post)["upper"],
    mean_f1_pre       = mean(f1_pre,       na.rm = TRUE),
    mean_f1_post      = mean(f1_post,      na.rm = TRUE),
    .groups = "drop"
  )
# NO_TREE cases remain in these summaries. A missing job is a fatal evaluation
# error, so every purity must contribute all 25 expected cases.
class_summary
write_csv(class_summary, file.path(out_dir, "classification_summary.csv"))

# --- Per-purity truncal / subclonal recall (Supplementary Fig. S8C) ---------
# These are the two components of the balanced accuracy above:
#   truncal recall (sensitivity)    = of the SVs whose truth is truncal, the
#                                     fraction the pipeline called truncal
#   subclonal recall (specificity)  = of the SVs whose truth is subclonal, the
#                                     fraction the pipeline called subclonal
# Fig. S8C previously carried these eight values hardcoded; it now reads this file.
recall_summary <- all_results %>%
  group_by(purity) %>%
  summarise(
    n                = n(),
    n_scored         = n(),
    n_no_tree        = sum(tree_outcome == "no_tree"),
    truncal_recall_pre    = mean(sens_pre,  na.rm = TRUE),
    subclonal_recall_pre  = mean(spec_pre,  na.rm = TRUE),
    truncal_recall_post   = mean(sens_post, na.rm = TRUE),
    subclonal_recall_post = mean(spec_post, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    bal_acc_pre_check  = (truncal_recall_pre  + subclonal_recall_pre)  / 2,
    bal_acc_post_check = (truncal_recall_post + subclonal_recall_post) / 2
  )
recall_summary
write_csv(recall_summary, file.path(out_dir, "recall_summary.csv"))
```

## Clone CCF MAE summary

```{r ccf-mae}
ccf_summary <- all_results %>%
  group_by(purity) %>%
  summarise(
    n_scored               = n(),
    n_no_tree              = sum(tree_outcome == "no_tree"),
    mean_mae_trunk_pre  = mean(ccf_mae_trunk_pre,  na.rm = TRUE),
    mean_mae_sub1_pre   = mean(ccf_mae_sub1_pre,   na.rm = TRUE),
    mean_mae_sub2_pre   = mean(ccf_mae_sub2_pre,   na.rm = TRUE),
    mean_mae_trunk_post = mean(ccf_mae_trunk_post,  na.rm = TRUE),
    mean_mae_sub1_post  = mean(ccf_mae_sub1_post,   na.rm = TRUE),
    mean_mae_sub2_post  = mean(ccf_mae_sub2_post,   na.rm = TRUE),
    .groups = "drop"
  )
ccf_summary
write_csv(ccf_summary, file.path(out_dir, "ccf_mae_summary.csv"))
```

## Tree topology recovery

```{r topology}
topo_summary <- all_results %>%
  group_by(purity) %>%
  summarise(
    n             = n(),
    n_scored      = n(),
    n_no_tree     = sum(tree_outcome == "no_tree"),
    n_correct     = sum(correct_topology, na.rm = TRUE),
    recovery_rate = n_correct / n,
    ci_lower      = wilson_ci(n_correct, n)["lower"],
    ci_upper      = wilson_ci(n_correct, n)["upper"],
    .groups = "drop"
  )
# NO_TREE is a scored topology failure. Missing output aborts the evaluation,
# preventing an incomplete denominator from being published.
topo_summary
write_csv(topo_summary, file.path(out_dir, "topology_summary.csv"))
```

## Table A2-1

```{r table-a2-1}
avg_ccf_mae <- all_results %>%
  group_by(purity) %>%
  summarise(
    pre_ccf_mae  = mean(c(ccf_mae_trunk_pre,  ccf_mae_sub1_pre,  ccf_mae_sub2_pre),  na.rm = TRUE),
    post_ccf_mae = mean(c(ccf_mae_trunk_post, ccf_mae_sub1_post, ccf_mae_sub2_post), na.rm = TRUE),
    .groups = "drop"
  )

table_a2_1 <- class_summary %>%
  left_join(avg_ccf_mae,  by = "purity") %>%
  left_join(topo_summary, by = "purity") %>%
  transmute(
    Purity             = paste0(purity, "%"),
    `N reps`           = n.x,
    `Pre bal. acc. [95% CI]`  = sprintf("%.3f [%.3f–%.3f]",
                                         mean_bal_acc_pre,  ci_lo_pre,  ci_hi_pre),
    `Post bal. acc. [95% CI]` = sprintf("%.3f [%.3f–%.3f]",
                                         mean_bal_acc_post, ci_lo_post, ci_hi_post),
    `Pre CCF MAE`      = sprintf("%.3f", pre_ccf_mae),
    `Post CCF MAE`     = sprintf("%.3f", post_ccf_mae),
    `Topology (n/N)`   = sprintf("%d/%d", n_correct, n.y)
  )

write_csv(table_a2_1, file.path(out_dir, "table_a2_1.csv"))
table_a2_1
```

## Table A2-2

```{r table-a2-2}
purities <- c(20L, 40L, 60L, 80L)

true_ccf_df <- do.call(rbind, lapply(purities, function(pur) {
  tc <- true_clone_fracs_s1(pur)
  data.frame(
    purity        = pur,
    clone         = c("trunk", "sub1", "sub2"),
    true_ccf_pre  = unname(tc$t1),
    true_ccf_post = unname(tc$t2),
    stringsAsFactors = FALSE
  )
}))

ccf_inf_long <- all_results %>%
  select(purity, starts_with("ccf_inf_")) %>%
  pivot_longer(cols = starts_with("ccf_inf_"),
               names_to = "metric", values_to = "ccf") %>%
  mutate(
    clone     = sub("ccf_inf_(trunk|sub1|sub2)_(pre|post)", "\\1", metric),
    timepoint = sub("ccf_inf_(trunk|sub1|sub2)_(pre|post)", "\\2", metric)
  )

table_a2_2 <- ccf_inf_long %>%
  group_by(purity, clone, timepoint) %>%
  summarise(
    mean_ccf = mean(ccf, na.rm = TRUE),
    sd_ccf   = sd(ccf,   na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  pivot_wider(names_from = timepoint,
              values_from = c(mean_ccf, sd_ccf),
              names_glue = "{timepoint}_{.value}") %>%
  left_join(true_ccf_df, by = c("purity", "clone")) %>%
  transmute(
    Purity = paste0(purity, "%"),
    Clone  = clone,
    `True CCF (pre)`              = sprintf("%.3f", true_ccf_pre),
    `Inferred CCF (pre, mean±SD)` = sprintf("%.3f±%.3f", pre_mean_ccf, pre_sd_ccf),
    `True CCF (post)`              = sprintf("%.3f", true_ccf_post),
    `Inferred CCF (post, mean±SD)` = sprintf("%.3f±%.3f", post_mean_ccf, post_sd_ccf)
  ) %>%
  arrange(Purity, Clone)

write_csv(table_a2_2, file.path(out_dir, "table_a2_2.csv"))
table_a2_2
```

## Figure A2-1

```{r fig-a2-1, fig.width=4, fig.height=9}
pub_theme <- theme_classic(base_size = 9) +
  theme(
    panel.grid.major = element_line(colour = "grey92", linewidth = 0.25),
    strip.background = element_rect(fill = "grey93", colour = NA),
    strip.text       = element_text(face = "bold", size = 9),
    axis.line        = element_line(linewidth = 0.35),
    axis.ticks       = element_line(linewidth = 0.35),
    axis.text        = element_text(colour = "black", size = 8),
    axis.title       = element_text(size = 9),
    legend.key.size  = unit(0.35, "cm"),
    legend.text      = element_text(size = 8),
    legend.title     = element_text(face = "bold", size = 9),
    plot.tag         = element_text(face = "bold", size = 10),
    plot.title       = element_text(size = 9)
  )
theme_set(pub_theme)

bal_acc_long <- class_summary %>%
  select(purity,
         Pre  = mean_bal_acc_pre,  pre_lo  = ci_lo_pre,  pre_hi  = ci_hi_pre,
         Post = mean_bal_acc_post, post_lo = ci_lo_post, post_hi = ci_hi_post) %>%
  pivot_longer(cols = c(Pre, Post),
               names_to = "timepoint", values_to = "mean_bal_acc") %>%
  mutate(
    ci_lo     = ifelse(timepoint == "Pre", pre_lo,  post_lo),
    ci_hi     = ifelse(timepoint == "Pre", pre_hi,  post_hi),
    timepoint = factor(timepoint, levels = c("Pre", "Post"))
  )

tp_colors <- c("Pre" = "#0072B2", "Post" = "#D55E00")

p_A <- bal_acc_long %>%
  ggplot(aes(x = purity, y = mean_bal_acc,
             color = timepoint, fill = timepoint, group = timepoint)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = purities, labels = paste0(purities, "%")) +
  scale_color_manual(values = tp_colors) +
  scale_fill_manual( values = tp_colors) +
  labs(x = NULL, y = "Balanced accuracy",
       color = "Timepoint", fill = "Timepoint") +
  scale_y_continuous(limits = c(0.5, 1), breaks = seq(0.5, 1, 0.1))

ccf_err_long <- all_results %>%
  select(purity, starts_with("ccf_err_")) %>%
  pivot_longer(cols = starts_with("ccf_err_"),
               names_to = "metric", values_to = "err") %>%
  mutate(
    clone     = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\1", metric),
    timepoint = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\2", metric),
    clone     = factor(clone, levels = c("trunk", "sub1", "sub2"),
                       labels = c("Trunk", "Subclone 1", "Subclone 2")),
    timepoint = factor(timepoint, levels = c("pre", "post"), labels = c("Pre", "Post"))
  ) %>%
  group_by(purity, clone, timepoint) %>%
  summarise(
    mean_err = mean(err, na.rm = TRUE),
    ci_lo    = normal_ci(err)["lower"],
    ci_hi    = normal_ci(err)["upper"],
    .groups  = "drop"
  )
write_csv(ccf_err_long, file.path(out_dir, "ccf_error_summary.csv"))

clone_colors <- c("Trunk" = "#555555", "Subclone 1" = "#009E73", "Subclone 2" = "#CC79A7")

p_B <- ccf_err_long %>%
  ggplot(aes(x = purity, y = mean_err,
             color = clone, fill = clone, linetype = timepoint,
             group = interaction(clone, timepoint))) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "grey50") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.10, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = purities, labels = paste0(purities, "%")) +
  scale_color_manual(values = clone_colors) +
  scale_fill_manual( values = clone_colors) +
  scale_linetype_manual(values = c("Pre" = "solid", "Post" = "dashed")) +
  labs(x = NULL, y = "CCF signed error (true − inferred)",
       color = "Clone", fill = "Clone", linetype = "Timepoint")

p_C <- topo_summary %>%
  ggplot(aes(x = factor(purity), y = recovery_rate)) +
  geom_col(fill = "#0072B2", width = 0.55) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                width = 0.18, linewidth = 0.5) +
  geom_text(aes(label = sprintf("%d/%d", n_correct, n),
                y = pmin(ci_upper + 0.04, 1.14)),
            size = 2.8) +
  scale_x_discrete(labels = paste0(purities, "%")) +
  labs(x = "Tumor purity", y = "Correct topology (fraction)") +
  coord_cartesian(ylim = c(0, 1.18))

p_a2_1 <- (p_A / p_B / p_C) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 10))

ggsave(file.path(out_dir, "fig_a2_1.pdf"), p_a2_1,
       width = 7, height = 9, useDingbats = FALSE)
p_a2_1
```

## Figure A2-2 (illustrative replicate, p = 60%)

```{r fig-a2-2, fig.width=7, fig.height=6}
illustrative_row <- all_results %>% filter(purity == 60L, tree_outcome == "tree") %>% slice(1)

if (nrow(illustrative_row) == 0) {
  message("No p=60% replicate found; skipping Figure A2-2.")
} else {
  boot_i <- illustrative_row$boot
  exp_i  <- illustrative_row$experiment

  longi_dir_i <- if (boot_i == 0L) {
    file.path(work_dir, "S1", paste0("c", cov, "p60"), exp_i)
  } else {
    file.path(work_dir, "S1", paste0("c", cov, "p60"),
              paste0("b", boot_i), exp_i)
  }

  sv2c_i  <- read_csv(file.path(longi_dir_i, "clustering", "sv2cluster.csv"),
                      show_col_types = FALSE)
  cents_i <- read_csv(file.path(longi_dir_i, "clustering", "cluster_centroids.csv"),
                      show_col_types = FALSE) %>%
    filter(pair == 1) %>%
    select(cluster_num, t1_ccf = f_pre, t2_ccf = f_day85)
  tree_i  <- readRDS(file.path(longi_dir_i, "tree", "tree_result.rds"))
  edges_i <- tree_i[[1]]

  trunk_idx_i   <- edges_i$child[edges_i$parent == "root"]
  trunk_clust_i <- if (length(trunk_idx_i) == 1)
    paste0("cluster", trunk_idx_i) else NA_character_

  sv2c_i <- sv2c_i %>%
    mutate(inferred_clonality = ifelse(cluster_num == trunk_clust_i,
                                       "truncal", "subclonal"))

  true_ccfs_60 <- true_clone_fracs_s1(60L)

  make_tree_plot <- function(node_df, edge_df, title_txt) {
    ggplot() +
      geom_segment(data = edge_df,
                   aes(x = x1, xend = x2, y = y1, yend = y2),
                   linewidth = 0.7, color = "grey40") +
      geom_label(data = node_df,
                 aes(x = x, y = y, label = label),
                 size = 2.4, fill = "white", label.size = 0.25,
                 label.padding = unit(0.25, "lines")) +
      xlim(-0.85, 0.85) + ylim(-0.4, 2.4) +
      theme_void(base_size = 9)
  }

  tree_edges_std <- data.frame(
    x1 = c(0,  0, 0), y1 = c(2, 1, 1),
    x2 = c(0, -0.6, 0.6), y2 = c(1, 0, 0)
  )

  true_nodes <- data.frame(
    label = c(
      "Root",
      "Trunk\nCCF = 1.00",
      sprintf("Sub1\npre=%.2f\npost=%.2f",
              true_ccfs_60$t1["sub1"], true_ccfs_60$t2["sub1"]),
      sprintf("Sub2\npre=%.2f\npost=%.2f",
              true_ccfs_60$t1["sub2"], true_ccfs_60$t2["sub2"])
    ),
    x = c(0, 0, -0.6, 0.6),
    y = c(2, 1, 0, 0)
  )
  p_A2 <- make_tree_plot(true_nodes, tree_edges_std, "")

  if (!is.null(edges_i) && nrow(edges_i) > 0 && !is.na(trunk_clust_i)) {
    non_trunk_i <- cents_i %>% filter(cluster_num != trunk_clust_i)
    inf_nodes <- data.frame(
      label = c(
        "Root",
        sprintf("Trunk\npre=%.2f\npost=%.2f",
                cents_i$t1_ccf[cents_i$cluster_num == trunk_clust_i],
                cents_i$t2_ccf[cents_i$cluster_num == trunk_clust_i]),
        if (nrow(non_trunk_i) >= 1)
          sprintf("Sub A\npre=%.2f\npost=%.2f",
                  non_trunk_i$t1_ccf[1], non_trunk_i$t2_ccf[1])
        else "Sub A\n(absent)",
        if (nrow(non_trunk_i) >= 2)
          sprintf("Sub B\npre=%.2f\npost=%.2f",
                  non_trunk_i$t1_ccf[2], non_trunk_i$t2_ccf[2])
        else "Sub B\n(absent)"
      ),
      x = c(0, 0, -0.6, 0.6),
      y = c(2, 1, 0, 0)
    )
    p_B2 <- make_tree_plot(inf_nodes, tree_edges_std, "")
  } else {
    p_B2 <- ggplot() +
      annotate("text", x = 0.5, y = 0.5,
               label = "Topology not\nrecovered", size = 3.5, hjust = 0.5) +
      theme_void()
  }

  scatter_df <- cents_i %>%
    arrange(cluster_num) %>%
    mutate(
      rank       = row_number(),
      is_trunk   = cluster_num == trunk_clust_i,
      sub_rank   = cumsum(!is_trunk),
      clone_label = case_when(
        is_trunk  ~ "Trunk",
        TRUE      ~ paste0("Sub ", sub_rank)
      )
    )

  p_C2 <- scatter_df %>%
    ggplot(aes(x = t1_ccf, y = t2_ccf, color = clone_label, label = clone_label)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey70",
                linewidth = 0.4) +
    geom_point(size = 3) +
    geom_text(vjust = -0.8, size = 2.8, show.legend = FALSE) +
    scale_color_manual(values = c("Trunk" = "#555555",
                                  "Sub 1" = "#009E73",
                                  "Sub 2" = "#CC79A7")) +
    labs(x = "Pre-treatment CCF", y = "Post-treatment CCF",
         color = "Clone") +
    coord_cartesian(xlim = c(0, 1.05), ylim = c(0, 1.05))

  truth_svs_i <- load_truth_beds(truth_dir, overlap = (exp_i != "exp1"))

  make_conf_df <- function(sv_tp, stage_label) {
    matched_tp <- sv_tp %>%
      rowwise() %>%
      mutate(
        truth_idx = {
          dists <- abs(truth_svs_i$start - POS)
          idx   <- which(truth_svs_i$chr == CHROM & dists < 500)
          if (length(idx) > 0) idx[which.min(dists[idx])] else NA_integer_
        },
        true_clone = if (!is.na(truth_idx)) truth_svs_i$true_clone[truth_idx] else NA_character_,
        true_clonality = case_when(
          true_clone == "truncal"            ~ "Truncal",
          true_clone %in% c("sub1", "sub2") ~ "Subclonal",
          TRUE                               ~ NA_character_
        )
      ) %>%
      ungroup() %>%
      filter(!is.na(true_clonality))

    if (nrow(matched_tp) == 0) return(NULL)

    matched_tp %>%
      count(true_clonality, inferred_clonality) %>%
      mutate(timepoint = stage_label)
  }

  pre_conf  <- make_conf_df(sv2c_i %>% filter(stage == pre_stage),  "Pre")
  post_conf <- make_conf_df(sv2c_i %>% filter(stage == post_stage), "Post")
  conf_df   <- bind_rows(pre_conf, post_conf)

  if (!is.null(conf_df) && nrow(conf_df) > 0) {
    p_D2 <- conf_df %>%
      mutate(
        inferred_clonality = tools::toTitleCase(inferred_clonality),
        inferred_clonality = factor(inferred_clonality, levels = c("Truncal", "Subclonal")),
        true_clonality     = factor(true_clonality, levels = c("Truncal", "Subclonal")),
        timepoint          = factor(timepoint, levels = c("Pre", "Post"))
      ) %>%
      ggplot(aes(x = inferred_clonality, y = true_clonality,
                 fill = n, label = n)) +
      geom_tile(color = "white", linewidth = 0.6) +
      geom_text(size = 3, color = "black") +
      facet_wrap(~timepoint) +
      scale_fill_gradient(low = "#e8f4f8", high = "#0072B2") +
      labs(x = "Inferred", y = "True", fill = "N SVs") +
      theme(panel.grid  = element_blank(),
            axis.line   = element_blank(),
            axis.ticks  = element_blank())
  } else {
    p_D2 <- ggplot() +
      annotate("text", x = 0.5, y = 0.5, label = "Insufficient data", size = 4) +
      theme_void()
  }

  p_a2_2 <- (p_A2 | p_B2) / (p_C2 | p_D2) +
    plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(face = "bold", size = 10))

  ggsave(file.path(out_dir, "fig_a2_2.pdf"), p_a2_2,
         width = 7, height = 6, useDingbats = FALSE)
  p_a2_2
}
```

## Multi-covariate correlation analysis (A2h)

```{r covariate-helpers}
build_covariate_long <- function(results_df, cov_prefix, log_transform = TRUE) {
  cov_long <- results_df %>%
    select(boot, purity, experiment,
           matches(paste0("^", cov_prefix,
                          "(trunk|sub1|sub2)_(pre|post)$"))) %>%
    pivot_longer(
      cols     = matches(paste0("^", cov_prefix,
                                "(trunk|sub1|sub2)_(pre|post)$")),
      names_to = "metric", values_to = "covariate"
    ) %>%
    mutate(
      clone     = sub(paste0(cov_prefix,
                             "(trunk|sub1|sub2)_(pre|post)"), "\\1", metric),
      timepoint = sub(paste0(cov_prefix,
                             "(trunk|sub1|sub2)_(pre|post)"), "\\2", metric)
    ) %>%
    select(-metric)

  err_long <- results_df %>%
    select(boot, purity, experiment,
           matches("^ccf_err_(trunk|sub1|sub2)_(pre|post)$")) %>%
    pivot_longer(
      cols     = matches("^ccf_err_(trunk|sub1|sub2)_(pre|post)$"),
      names_to = "metric", values_to = "signed_error"
    ) %>%
    mutate(
      clone     = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\1", metric),
      timepoint = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\2", metric)
    ) %>%
    select(-metric)

  cov_long %>%
    left_join(err_long,
              by = c("boot", "purity", "experiment", "clone", "timepoint")) %>%
    filter(!is.na(signed_error), !is.na(covariate),
           if (log_transform) covariate > 0 else TRUE) %>%
    mutate(
      covariate_val = if (log_transform) log10(covariate) else covariate,
      replicate     = paste0("b", boot, "_p", purity, "_", experiment),
      clone         = factor(clone,
                             levels = c("trunk", "sub1", "sub2"),
                             labels = c("Trunk", "Subclone 1", "Subclone 2")),
      timepoint     = factor(timepoint,
                             levels = c("pre", "post"),
                             labels = c("Pre", "Post"))
    )
}

run_covariate_analysis <- function(df_in, label, n_boot_ci = 1000L) {
  clone_levels <- levels(df_in$clone)

  spearman_results <- lapply(clone_levels, function(cl) {
    sub_df  <- df_in %>% filter(clone == cl)
    rep_ids <- unique(sub_df$replicate)
    if (nrow(sub_df) < 3L) return(NULL)
    obs_rho <- cor(sub_df$covariate_val, sub_df$signed_error,
                   method = "spearman", use = "complete.obs")
    boot_rhos <- replicate(n_boot_ci, {
      sampled <- sample(rep_ids, length(rep_ids), replace = TRUE)
      bs_df   <- bind_rows(lapply(sampled,
                                  function(r) sub_df[sub_df$replicate == r, ]))
      cor(bs_df$covariate_val, bs_df$signed_error,
          method = "spearman", use = "complete.obs")
    })
    ci <- quantile(boot_rhos, c(0.025, 0.975), na.rm = TRUE)
    data.frame(clone = cl, rho = obs_rho,
               ci_lo = ci[[1]], ci_hi = ci[[2]],
               n_obs = nrow(sub_df), n_reps = length(rep_ids),
               stringsAsFactors = FALSE)
  })
  spearman_df <- bind_rows(Filter(Negate(is.null), spearman_results))

  r2_df <- bind_rows(lapply(clone_levels, function(cl) {
    sub_df <- df_in %>% filter(clone == cl)
    if (nrow(sub_df) < 3L) return(NULL)
    fit <- lm(signed_error ~ covariate_val, data = sub_df)
    data.frame(clone = cl,
               r_squared = summary(fit)$r.squared,
               stringsAsFactors = FALSE)
  }))

  lrt_pval <- NA_real_
  m_full <- tryCatch(
    lmer(signed_error ~ covariate_val * clone + (1 | replicate),
         data = df_in, REML = FALSE),
    error = function(e) { cat("  lmer (full) failed:", conditionMessage(e), "\n"); NULL }
  )
  m_null <- tryCatch(
    lmer(signed_error ~ covariate_val + clone + (1 | replicate),
         data = df_in, REML = FALSE),
    error = function(e) { cat("  lmer (null) failed:", conditionMessage(e), "\n"); NULL }
  )
  if (!is.null(m_full) && !is.null(m_null)) {
    lrt_res  <- anova(m_null, m_full)
    lrt_pval <- lrt_res[["Pr(>Chisq)"]][2]
    coefs    <- as.data.frame(coef(summary(m_full)))
    coefs$term <- rownames(coefs)
    cat(sprintf("  Interaction LRT p = %.4g\n", lrt_pval))
    write_csv(coefs,
      file.path(out_dir, sprintf("mixed_model_coefs_%s.csv", label)))
  }

  list(spearman_df = spearman_df, r2_df = r2_df, lrt_pval = lrt_pval)
}

make_covariate_fig <- function(df_in, spearman_df, r2_df, x_label, title) {
  clone_colors <- c("Trunk" = "#555555", "Subclone 1" = "#009E73", "Subclone 2" = "#CC79A7")
  rho_labels <- spearman_df %>%
    mutate(clone = factor(clone, levels = c("Trunk", "Subclone 1", "Subclone 2"))) %>%
    left_join(r2_df, by = "clone") %>%
    mutate(label = sprintf(
      "atop(rho == %.2f~'['*%.2f*','~%.2f*']',~R^2 == %.3f)",
      rho, ci_lo, ci_hi, r_squared)) %>%
    left_join(
      df_in %>% group_by(clone) %>%
        summarise(x_pos = min(covariate_val, na.rm = TRUE), .groups = "drop"),
      by = "clone"
    )

  df_in %>%
    ggplot(aes(x = covariate_val, y = signed_error,
               color = clone, fill = clone)) +
    geom_hline(yintercept = 0, linetype = "dotted", color = "grey50") +
    geom_point(alpha = 0.40, size = 1.6, shape = 16) +
    geom_smooth(method = "lm", formula = y ~ x, se = TRUE,
                alpha = 0.15, linewidth = 0.9) +
    geom_text(data = rho_labels,
              aes(x = x_pos, y = Inf, label = label),
              hjust = 0, vjust = 1.4, size = 3.2, color = "black",
              parse = TRUE, inherit.aes = FALSE) +
    scale_color_manual(values = clone_colors) +
    scale_fill_manual( values = clone_colors) +
    facet_wrap(~clone, scales = "free_x", ncol = 1) +
    labs(x = x_label, y = "Signed CCF error (true − inferred)",
         caption = sprintf("n = %d (replicate × clone × timepoint).", nrow(df_in))) +
    theme(legend.position = "none",
          plot.caption    = element_text(hjust = 0, size = 7, colour = "grey40"))
}
```

```{r covariate-analysis, fig.width=4, fig.height=8}
set.seed(42L)

covariates <- list(
  list(prefix        = "eff_cov_",
       label         = "ao",
       display_label = "Mean AO (SVtyper)",
       x_label       = expression(log[10](mean~AO~per~clone)),
       log_transform = TRUE),
  list(prefix        = "bpc_",
       label         = "bpc",
       display_label = "Number of SVs assigned to clone",
       x_label       = expression(log[10](n~SVs~assigned~to~clone)),
       log_transform = TRUE),
  list(prefix        = "seg_len_",
       label         = "seg_len",
       display_label = "Mean SV segment length",
       x_label       = expression(log[10](mean~SV~length~(bp))),
       log_transform = TRUE),
  list(prefix        = "ascn_",
       label         = "ascn",
       display_label = "ASCN mismatch fraction",
       x_label       = "ASCN mismatch fraction (non-diploid)",
       log_transform = FALSE)
)

all_cov_results <- lapply(covariates, function(cov) {
  cat(sprintf("\n--- Covariate: %s ---\n", cov$display_label))
  results_for_cov <- if (cov$label == "bpc") {
    all_results %>%
      mutate(
        bpc_trunk_pre  = bpc_trunk_pre + bpc_trunk_post,
        bpc_sub1_pre   = bpc_sub1_pre  + bpc_sub1_post,
        bpc_sub2_pre   = bpc_sub2_pre  + bpc_sub2_post,
        bpc_trunk_post = bpc_trunk_pre,
        bpc_sub1_post  = bpc_sub1_pre,
        bpc_sub2_post  = bpc_sub2_pre
      )
  } else all_results
  df_c <- build_covariate_long(results_for_cov, cov$prefix, cov$log_transform)
  if (cov$label == "bpc") {
    df_c <- df_c %>%
      group_by(boot, purity, experiment, clone, replicate,
               covariate, covariate_val) %>%
      summarise(signed_error = mean(signed_error, na.rm = TRUE),
                .groups = "drop")
  }
  cat(sprintf("  n = %d observations\n", nrow(df_c)))
  if (nrow(df_c) < 10L) { cat("  Insufficient data, skipping.\n"); return(NULL) }

  write_csv(df_c, file.path(out_dir, sprintf("covariate_data_%s.csv", cov$label)))

  res <- run_covariate_analysis(df_c, cov$label)
  cat("  Spearman rho [95% CI]:\n")
  print(res$spearman_df, digits = 3)

  spearman_with_r2 <- res$spearman_df %>% left_join(res$r2_df, by = "clone")
  write_csv(spearman_with_r2, file.path(out_dir, sprintf("spearman_%s.csv", cov$label)))

  write_csv(
    data.frame(covariate = cov$label, lrt_pval = res$lrt_pval),
    file.path(out_dir, sprintf("lrt_pval_%s.csv", cov$label))
  )

  p <- make_covariate_fig(df_c, res$spearman_df, res$r2_df, cov$x_label,
                          sprintf("Signed CCF error vs. %s", cov$display_label))
  out_pdf <- file.path(out_dir, sprintf("fig_a2_3_%s.pdf", cov$label))
  ggsave(out_pdf, p, width = 4, height = 8, useDingbats = FALSE)
  print(p)

  c(list(df = df_c, label = cov$label, display_label = cov$display_label), res)
})
names(all_cov_results) <- vapply(covariates, `[[`, character(1), "label")
```

## Spearman summary (all covariates)

```{r spearman-summary}
spearman_rows <- bind_rows(lapply(all_cov_results, function(r) {
  if (is.null(r)) return(NULL)
  r$spearman_df %>%
    left_join(r$r2_df, by = "clone") %>%
    mutate(covariate = r$display_label, lrt_pval = r$lrt_pval)
}))
if (nrow(spearman_rows)) {
  spearman_summary <- spearman_rows %>%
    select(covariate, clone, rho, ci_lo, ci_hi, r_squared, lrt_pval, n_obs)
} else {
  spearman_summary <- tibble(
    covariate = character(), clone = character(), rho = double(),
    ci_lo = double(), ci_hi = double(), r_squared = double(),
    lrt_pval = double(), n_obs = integer()
  )
}

spearman_summary
write_csv(spearman_summary, file.path(out_dir, "spearman_summary_all.csv"))

ao_res <- all_cov_results[["ao"]]
if (!is.null(ao_res)) {
  write_csv(ao_res$df,        file.path(out_dir, "coverage_correlation_data.csv"))
  write_csv(ao_res$spearman_df, file.path(out_dir, "spearman_results.csv"))
  file.copy(file.path(out_dir, "fig_a2_3_bpc.pdf"),
            file.path(out_dir, "fig_a2_3.pdf"), overwrite = TRUE)
}
```

## SV-type stratified analysis

```{r sv-type-stratified, fig.width=7, fig.height=4}
if (!is.null(sv_detail_df) && nrow(sv_detail_df) > 0L) {

  err_ref <- all_results %>%
    select(boot, purity, experiment,
           matches("^ccf_err_(trunk|sub1|sub2)_(pre|post)$")) %>%
    pivot_longer(
      cols     = matches("^ccf_err_(trunk|sub1|sub2)_(pre|post)$"),
      names_to = "metric", values_to = "signed_error"
    ) %>%
    mutate(
      clone     = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\1", metric),
      timepoint = sub("ccf_err_(trunk|sub1|sub2)_(pre|post)", "\\2", metric),
      replicate = paste0("b", boot, "_p", purity, "_", experiment),
      clone     = factor(clone, levels = c("trunk", "sub1", "sub2"),
                         labels = c("Trunk", "Subclone 1", "Subclone 2")),
      timepoint = factor(timepoint, levels = c("pre", "post"),
                         labels = c("Pre", "Post"))
    ) %>%
    select(-metric) %>%
    filter(!is.na(signed_error))

  sv_labeled <- sv_detail_df %>%
    mutate(
      replicate = paste0("b", boot, "_p", purity, "_", experiment),
      clone     = factor(clone, levels = c("trunk", "sub1", "sub2"),
                         labels = c("Trunk", "Subclone 1", "Subclone 2")),
      timepoint = factor(timepoint, levels = c("pre", "post"),
                         labels = c("Pre", "Post")),
      sv_type   = toupper(sv_type)
    ) %>%
    left_join(err_ref, by = c("replicate", "clone", "timepoint",
                               "boot", "purity", "experiment"))

  sv_agg <- sv_labeled %>%
    group_by(replicate, clone, timepoint, sv_type) %>%
    summarise(
      mean_ao      = mean(ao,            na.rm = TRUE),
      bpc_count    = n(),
      mean_seg_len = mean(sv_len,        na.rm = TRUE),
      ascn_frac    = mean(is_nondiploid, na.rm = TRUE),
      signed_error = first(signed_error),
      .groups = "drop"
    ) %>%
    filter(!is.na(signed_error))

  write_csv(sv_agg, file.path(out_dir, "sv_type_stratified_data.csv"))

  sv_cov_defs <- list(
    list(col = "mean_ao",      log = TRUE,  label = "Mean AO"),
    list(col = "bpc_count",    log = TRUE,  label = "N SVs assigned"),
    list(col = "mean_seg_len", log = TRUE,  label = "Mean seg length"),
    list(col = "ascn_frac",    log = FALSE, label = "ASCN mismatch")
  )

  strat_rho_rows <- list()
  for (cov_def in sv_cov_defs) {
    sv_sub <- sv_agg %>%
      rename(covariate = !!sym(cov_def$col)) %>%
      filter(!is.na(covariate),
             if (cov_def$log) covariate > 0 else TRUE) %>%
      mutate(covariate_val = if (cov_def$log) log10(covariate) else covariate)

    for (cl in c("Trunk", "Subclone 1", "Subclone 2")) {
      for (svt in unique(sv_sub$sv_type)) {
        sub_df <- sv_sub %>% filter(clone == cl, sv_type == svt)
        if (nrow(sub_df) < 5L) next
        rho <- cor(sub_df$covariate_val, sub_df$signed_error,
                   method = "spearman", use = "complete.obs")
        strat_rho_rows <- c(strat_rho_rows, list(data.frame(
          covariate = cov_def$label, clone = cl, sv_type = svt,
          rho = rho, n = nrow(sub_df), stringsAsFactors = FALSE
        )))
      }
    }
  }

  strat_rho_df <- bind_rows(strat_rho_rows)
  strat_rho_df
  write_csv(strat_rho_df, file.path(out_dir, "spearman_sv_type_stratified.csv"))

  if (nrow(strat_rho_df) > 0L) {
    p_heatmap <- strat_rho_df %>%
      mutate(clone     = factor(clone, levels = c("Trunk", "Subclone 1", "Subclone 2")),
             sv_type   = factor(sv_type),
             covariate = factor(covariate,
                                levels = c("Mean AO", "N SVs assigned",
                                           "Mean seg length", "ASCN mismatch"))) %>%
      ggplot(aes(x = sv_type, y = covariate, fill = rho)) +
      geom_tile(color = "white", linewidth = 0.5) +
      geom_text(aes(label = sprintf("%.2f\n(n=%d)", rho, n)),
                size = 2.5, lineheight = 0.9) +
      scale_fill_gradient2(low = "#0072B2", mid = "white", high = "#D55E00",
                           midpoint = 0, limits = c(-1, 1),
                           name = expression(Spearman~rho)) +
      facet_wrap(~clone, ncol = 3) +
      labs(x = "SV type", y = NULL) +
      theme(axis.text.x = element_text(angle = 30, hjust = 1),
            panel.grid  = element_blank(),
            axis.line   = element_blank(),
            axis.ticks  = element_blank())

    ggsave(file.path(out_dir, "fig_a2_3_sv_type_heatmap.pdf"),
           p_heatmap, width = 7, height = 4, useDingbats = FALSE)
    p_heatmap
  }

} else {
  message("No SV detail data available for SV-type stratification.")
}
```

## Decision rules (AO covariate)

```{r decision-rules}
if (!is.null(ao_res)) {
  for (cl in c("Subclone 1", "Subclone 2", "Trunk")) {
    r <- ao_res$spearman_df[ao_res$spearman_df$clone == cl, ]
    if (nrow(r) == 0L) { cat(sprintf("  %s: no data\n", cl)); next }
    if (cl != "Trunk") {
      met <- r$rho < -0.3 && r$ci_hi < 0
      cat(sprintf(
        "  %s: rho = %+.3f [%+.3f, %+.3f]  rho < -0.3 and CI excludes 0: %s\n",
        cl, r$rho, r$ci_lo, r$ci_hi, if (met) "YES" else "NO"))
    } else {
      met <- abs(r$rho) < 0.2 && r$ci_lo < 0 && r$ci_hi > 0
      cat(sprintf(
        "  Trunk: rho = %+.3f [%+.3f, %+.3f]  |rho| < 0.2 and CI includes 0: %s\n",
        r$rho, r$ci_lo, r$ci_hi, if (met) "YES" else "NO"))
    }
  }
  cat(sprintf("  Interaction LRT (AO): p = %.4g  p < 0.05: %s\n",
              ao_res$lrt_pval,
              if (!is.na(ao_res$lrt_pval) && ao_res$lrt_pval < 0.05) "YES" else "NO"))
}
```
